---
title: "grid plot"
author: "Jen Richmond"
date: "12/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(janitor)
```


## R Markdown

```{r}
aqi <- read_csv(here::here("data", "nswgov_daily.csv")) %>%
  clean_names()
```

```{r}
aqi$date <- dmy(aqi$date)

glimpse(aqi)
```

# Make it long
```{r}
aqi_long <- aqi %>%
  select(1:27) %>%
  pivot_longer(names_to = "site", values_to = "aqi", 2:27)
```

select just sydney
```{r}
sydney <- aqi_long %>%
  filter(site == "sydney_central_east_raqi_24_hour_index") 

```

# just 2019

this is not what we want, percentiles over 2019, not that date across years

```{r}
sydney2019 <- sydney %>%
  filter(date > "2018-12-31")

percentile <- sydney2019 %>%
  mutate(percentile = percent_rank(aqi)) %>%
  mutate(percent_100 = percentile*100) 
```

# new approach

use yday function from lubriate to get a day number  
```{r}
sydney <- sydney %>%
  mutate(day = yday(date), month = month(date), year = year(date))  

sydney$month <- as.factor(sydney$month)
```

```{r}
glimpse(sydney)
```

This added percentile by day, but there are only 5 values. So maybe better to look at today relative to a month. 

```{r}
day_percent  <- sydney %>%
  group_by(day) %>%
  mutate(percent = percent_rank(aqi)*100)
```

adding percentile column with percentiles by month
```{r}
month_percent <- sydney %>%
  group_by(month) %>%
  mutate(percent = percent_rank(aqi)*100)

#write to csv

month_percent %>%
  write_csv(here("data", "aqi_percentile_month.csv"))
```

plot colour percent 
```{r}
month_percent %>%
  ggplot(aes(x = month, y = aqi, colour = percent)) +
  geom_point() +
  scale_color_gradient(low="blue", high="red") +
  facet_wrap(~ year) +
  theme_classic()

```

```{r}
# Library
library(ggplot2)

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)
 
# Heatmap 
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile()
```

We want y = month, x = day z = smokiness, but we will need to create dates that dont exist. 
```{r}

```

