---
title: "grid plot"
author: "Jen Richmond"
date: "12/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmd reads in the daily average AQI and calculates the percentile for each daily value relative to that month. Writes percentile values to csv file = aqi_percentile_month.csv

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```


## R Markdown

```{r}
aqi <- read_csv(here::here("data", "nswgov_daily.csv")) %>%
  clean_names()
```

```{r}
aqi$date <- dmy(aqi$date)

glimpse(aqi)
```

# Make it long
```{r}
aqi_long <- aqi %>%
  select(1:27) %>%
  pivot_longer(names_to = "site", values_to = "aqi", 2:27)
```

select just sydney
```{r}
sydney <- aqi_long %>%
  filter(site == "sydney_central_east_raqi_24_hour_index") 

```

# Just Sydney_2019

this is not what we want, percentiles over 2019, not that date across years
```{r}
sydney2019 <- sydney %>%
  filter(date > "2018-12-31")

percentile <- sydney2019 %>%
  mutate(percentile = percent_rank(aqi)) %>%
  mutate(percent_100 = percentile*100) 
```

# Percentiles 

use yday function from lubriate to get a day number  
```{r}
sydney <- sydney %>%
  mutate(day = yday(date), month = month(date), year = year(date))  

sydney$month <- as.factor(sydney$month)
```

### By day number 
This added percentile by day, but there are only 5 values. So maybe better to look at today relative to a month. 

```{r}
day_percent  <- sydney %>%
  group_by(day) %>%
  mutate(percent = percent_rank(aqi)*100)
```

### By month 
```{r}
month_percent <- sydney %>%
  group_by(month) %>%
  mutate(percent = percent_rank(aqi)*100)

#write to csv

month_percent %>%
  write_csv(here::here("data", "aqi_percentile_month.csv"))
```


